### 2019-04-01 paper-template.Rmd
@Fsolt original template.

---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    latex_engine: pdflatex
    template: svm-latex-ms2.tex
title: "An Example Article" 
thanks: "The paper's revision history and the materials needed to reproduce its analyses can be found [on Github here](http://github.com/fsolt/example_article). Corresponding author: [frederick-solt@uiowa.edu](mailto:frederick-solt@uiowa.edu). Current version: `r format(Sys.time(), '%B %d, %Y')`."
author:
- name: Ahmad S. Qabazard
  affiliation: University of Iowa
abstract: "Here's where you write 100 to 250 words, depending on the journal, that describe your objective, methods, results, and conclusion."
keywords: "these, always seem silly, to me, given google, but regardless"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
spacing: single
bibliography: \dummy{`r file.path(getwd(), list.files(getwd(), "bib$"))`}
biblio-style: apsr
citecolor: black
linkcolor: black
endnote: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# load all the packages you will use below 
library(dotwhisker)
library(tidyverse)
```

# Introduction to RMarkdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r mtcars}
summary(mtcars)
```

---
# This is a comment, set off with --- and started with #.  Comments are good for notes to self that you don't want to show up in the output.  Below is LaTeX code for a page break.
---
\pagebreak 
## Including Plots

You can also embed plots, for example:

```{r dotwhisker_plot, fig.width = 7, fig.height = 4, warning = FALSE, message = FALSE, echo = FALSE, fig.cap="\\label{fig:dotwhisker_plot}Dot-and-Whisker Plot Example"}
m1 <- lm(mpg ~ wt + cyl + disp + gear, data = mtcars)
m2 <- update(m1, . ~ . + hp) # add another predictor
m3 <- update(m2, . ~ . + am) # and another 
dwplot(list(m1, m2, m3),
       vline = geom_vline(xintercept = 0, colour = "grey60", linetype = 2)) %>% # plot line at zero _behind_ coefs
    relabel_predictors(c(wt = "Weight",                       
                         cyl = "Cylinders", 
                         disp = "Displacement", 
                         hp = "Horsepower", 
                         gear = "Gears", 
                         am = "Manual")) +
     theme_bw() + xlab("Coefficient Estimate") + ylab("") +
     geom_vline(xintercept = 0, colour = "grey60", linetype = 2) +
     ggtitle("Predicting Gas Mileage") +
     theme(plot.title = element_text(face="bold"),
           legend.position = c(0.007, 0.01),
           legend.justification = c(0, 0), 
           legend.background = element_rect(colour="grey80"),
           legend.title = element_blank()) 
```

Figure \ref{fig:dotwhisker_plot} is a plot made using the \texttt{dotwhisker} package [@Solt2015c]. Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Citations
Want to cite something?

1. Find your cite key in your bib file.
1. Put an @ before it, like @Solt2017, or whatever it is
1. @Solt2017 creates an in-text citation
1. [@Herndon2014] creates a parenthetical citation

As @Gelman2014 note, the garden of forking paths can pose problems for researchers even when they are acting in good faith.

#Zotero

## Other Common Things

> This will create a block quote, if you want one.
Dropping a footnote is easy.^[See? Not hard at all.]